# Combine abundance information  

Pull together data from all life stages. There are a couple of graphs in here; life stages are not in order and I did not try to make them look nice, but they will give a very rough idea of what's in the data.    

```{r}
library(tidyverse)
library(plotly)

source(here::here("functions.R"))

in_path <- here::here("data", 
                      "processed",
                      "byLifeStage")

in_env <- here::here("data",
                     "processed",
                     "environmental")

out_path <- here::here("data",
                       "processed")
```


The files for each life stage have the same columns in them. Some have additional columns that are removed at this stage.  

```{r}
in_files <- paste0(
     c("postlarval",
      "juvenile",
      "subadult",
      "adult",
      "adult.landings.brown",
      "adult.landings.white"),
    "_byShrimpYear.csv")

in_paths <- file.path(in_path, in_files)
```

I will create a few files: one with all abundance information for both species, so future data users can compare whichever subsets they are interested in. That will be in a subfolder `data/processed/combinedLifeStages`.  

Then I will split brown and white shrimp, and combine the environmental data with abundance data. These will be saved in `data/processed`.  

Landings data are retained in these files. Trip Ticket data has not been incorporated.  

```{r}
dat <- in_paths |> 
    purrr::map(read.csv) |> 
    dplyr::bind_rows() |> 
    select(all_of(sel_order_annual),
           abundanceIndex_fullTimePeriod,
           abundanceIndex_commonTimePeriod)
```


```{r}
write.csv(dat,
          here::here(out_path,
                     "combinedLifeStages",
                     "allAbundanceIndices.csv"),
          row.names = FALSE,
          na = "")
```

## Brown Shrimp  

```{r}
brn <- dat |> 
    filter(species == "brown")

brn_env <- read.csv(here::here(in_env,
                               "brn_environmental.csv")) |> 
    filter(shrimp_year %in% unique(brn$shrimp_year))

brn_combined <- left_join(brn, brn_env)

```

Write it out.  

```{r}
write.csv(brn_combined,
          here::here(out_path,
                     "brn_abundAndEnv.csv"),
          na = "",
          row.names = FALSE)
```


```{r}

p <- ggplot(brn,
       aes(x = shrimp_year,
           y = abundanceIndex_fullTimePeriod,
           col = life_stage)) +
    geom_hline(yintercept = 0, col = "gray40") +
    geom_line(linewidth = 0.5,
              alpha = 0.4) +
    geom_point() +
    theme_bw() +
    labs(title = "Brown Shrimp",
         x = "Shrimp Year",
         y = "Abundance Index",
         col = "Life Stage") +
    theme(legend.position = "bottom")

plotly_out(p)
```



## White Shrimp  

```{r}
wht <- dat |> 
    filter(species == "white")

wht_env <- read.csv(here::here(in_env,
                               "wht_environmental.csv")) |> 
    filter(shrimp_year %in% unique(wht$shrimp_year))

wht_combined <- left_join(wht, wht_env)

write.csv(wht_combined,
          here::here(out_path,
                     "wht_abundAndEnv.csv"),
          na = "",
          row.names = FALSE)
```

```{r}

p <- ggplot(wht,
       aes(x = shrimp_year,
           y = abundanceIndex_fullTimePeriod,
           col = life_stage)) +
    geom_hline(yintercept = 0, col = "gray40") +
    geom_line(linewidth = 0.5,
              alpha = 0.4) +
    geom_point() +
    theme_bw() +
    labs(title = "White Shrimp",
         x = "Shrimp Year",
         y = "Abundance Index",
         col = "Life Stage") +
    theme(legend.position = "bottom")

plotly_out(p)

```




## About

**Input `.qmd` file** for this section was: **`r xfun::with_ext(knitr::current_input(), "qmd")`** and can be found in the main directory.  

**Input data file(s)** - in `data/processed/byLifeStage`:  

-   `postlarval_byShrimpYear.csv`
-   `juvenile_byShrimpYear.csv`
-   `subadult_byShrimpYear.csv`  
-   `adult_byShrimpYear.csv`  
-   `adult.landings.brown_byShrimpYear.csv`  
-   `adult.landings.white_byShrimpYear.csv`  


**Modifications to data:**  

-   reduced to only common shrimp abundance columns  
-   combined abundance and environmental data, based on shrimp year.  
-   **NOTE:** because the abundances are in a long format, meaning each life stage has a separate row for shrimp year, the same environmental data is repeated on multiple rows.


**Generated data file(s):**   

-   `allAbundanceIndices.csv`, in `data/processed/combinedLifeStages`. Contains abundance indices for each life stage and shrimp year, for both brown and white shrimp.  
-   `brn_abundAndEnv.csv`, in `data/processed`. Contains abundance index and environmental metrics (nursery period salinity; winter days below temperature thresholds) for brown shrimp only.  
-   `wht_abundAndEnv.csv`, in `data/processed`. Contains abundance index and environmental metrics (nursery period salinity; winter days below temperature thresholds) for white shrimp only.  


**Other notes about data:**

- none