# Adults - landings and trip tickets  

#### Setup 

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

source(here::here("functions.R"))

in_path <- here::here("data", "intermediate")
out_path <- here::here("data", "processed", "byLifeStage")
out_path_detailed <- here::here(out_path, "moreDetailed")

load(here::here(in_path, "adultCommercial_dfs.RData"))
```

## Landings Data  

### Brown Shrimp  

Use annual data for brown shrimp, because it corresponds to the life cycle and shrimp year.  Three states are represented in this dataset. NC is not represented elsewhere, so it is removed and we will focus only on GA and SC. 

```{r}
brn_landings <- landings_annual |> 
    janitor::clean_names() |> 
    filter(species == "brown",
           state != "NORTH CAROLINA") |> 
    select(shrimp_year = year,
           species,
           state,
           pounds)

brn_abund <- brn_landings |> 
    mutate(survey = case_when(state == "GEORGIA" ~ "GA",
                              state == "SOUTH CAROLINA" ~ "SC",
                              .default = "PROBLEM"),
           life_stage = "adult.landings",
           abundance_measure = "pounds",
           abundance = pounds,
           sqrt_abundance = sqrt(abundance)) |> 
    select(all_of(sel_order_annual)) |> 
    arrange(shrimp_year, species, survey)
```

```{r}
#| label: fig-brnAbundByState
#| fig-caption: "Brown shrimp annual landings by state"

ggplot(brn_abund,
       aes(x = shrimp_year,
           y = abundance,
           col = survey)) +
    geom_point(size = 2)
```

Total over the two states (because the data we have for white shrimp is totals across the states, not averages) and generate abundance index.  

```{r}
brn_byShrimpYear <- brn_abund |> 
    group_by(species, life_stage, shrimp_year) |> 
    summarize(abundance = sum(abundance, na.rm = TRUE)) |> 
    ungroup() |> 
    mutate(sqrt_abundance = sqrt(abundance),
           survey = "Landings, GA and SC",
           abundance_measure = "total over 2 states") |> 
    select(all_of(sel_order_annual)) |> 
    group_by(species) |> 
    mutate(abundanceIndex_fullTimePeriod = as.vector(scale(sqrt_abundance))) |> 
    ungroup()

brn_abund_commontime <- brn_byShrimpYear |> 
    select(species, shrimp_year, sqrt_abundance) |> 
    filter(shrimp_year %in% common_shrimpYears) |> 
    group_by(species) |> 
    mutate(abundanceIndex_commonTimePeriod = as.vector(scale(sqrt_abundance))) |> 
    ungroup()

brn_byShrimpYear <- left_join(brn_byShrimpYear, 
                              brn_abund_commontime)
```

Write out.  

```{r}
# by state
write.csv(brn_abund,
          here::here(out_path_detailed,
                     "adult.landings.brown_byStateAndShrimpYear.csv"),
          na = "",
          row.names = FALSE)

# combined with index
write.csv(brn_byShrimpYear,
          here::here(out_path,
                     "adult.landings.brown_byShrimpYear.csv"),
          na = "",
          row.names = FALSE)
```


### White Shrimp  

Using seasonal data for white shrimp, to separate non-spawning adults (time period 2 for a calendar year, season Fall, Months Aug-Dec) from spawners (time period 1 of the next calendar year, season Winter-Spring-Summer, months Jan-July).   

The input file contained heads-off landings from SC and GA.  

```{r}
# Fall = Aug-Dec  - make shrimp_year =  year
# Win_Spr_Sum = Jan-Jul - make shrimp_year = year - 1

wht_landings <- landings_seas |> 
    mutate(shrimp_year = case_when(season == "Fall" ~ Year,
                                   .default = Year - 1),
           life_stage = case_when(season == "Fall" ~ "adult.nonspawning.landings",
                                  .default = "adult.spawning.landings"),
           survey = "landings",
           species = "white",
           abundance_measure = "head-off weight, sum of SC and GA",
           abundance = lbs_headOff,
           sqrt_abundance = sqrt(abundance)) |> 
    filter(shrimp_year >= 1979,
           shrimp_year <= 2022) |> 
    janitor::clean_names() |> 
    select(any_of(sel_order_monthly),
           season, sqrt_abundance) |> 
    relocate(season, .after = "year")
```

Generate abundance index. Handling a little differently because there are two life stages in here (spawning and non-spawning adults).  

```{r}
# abundance index, full time
wht_abund <- wht_landings |> 
    group_by(species, life_stage) |> 
    mutate(abundanceIndex_fullTimePeriod = as.vector(scale(sqrt_abundance))) |> 
    ungroup()

# abundance index, common time
wht_abund_commonTime <- wht_abund |> 
    select(species, life_stage, shrimp_year, sqrt_abundance) |> 
    filter(shrimp_year %in% common_shrimpYears) |> 
    group_by(species, life_stage) |> 
    mutate(abundanceIndex_commonTimePeriod = as.vector(scale(sqrt_abundance))) |> 
    ungroup()

wht_abund <- left_join(wht_abund, 
                       wht_abund_commonTime) |> 
    rename(calendar_year = year) |> 
    select(survey, species, shrimp_year, life_stage,
           calendar_year, season,
           everything())
```

```{r}
#| label: fig-whtAbundIndex
#| fig-cap: "Abundance Index for white shrimp, by spawning vs. non-spawning adults"


ggplot(wht_abund,
       aes(x = shrimp_year,
           y = abundanceIndex_fullTimePeriod,
           col = life_stage)) +
    geom_point(size = 2) +
    geom_line() +
    theme_bw()
```


```{r}
#| eval: false 

ggplot(wht_abund) +
    geom_histogram(aes(x = abundanceIndex_fullTimePeriod)) +
    facet_wrap(~life_stage)
```

Write out white shrimp data. Both 'adult' life stages are in the single file.  

```{r}
write.csv(wht_abund,
          here::here(out_path,
                     "adult.landings.white_byShrimpYear.csv"),
          na = "",
          row.names = FALSE)
```



## Trip Ticket Data  

For browns, compile to annual totals.  

For whites, ideally compile to the same seasons reported in landings: Jan-Jul and Aug-Dec. However the waves here split; Jul-Aug are combined. Based on the nursery period dates for this species starting in June, I'll ..... cut it off at June?

```{r}
#| label: fig-WaveComparison
#| fig-cap: "White Shrimp Landings in GA and SC, by wave. These landings are reported as head-off weights (pounds) by state. They were not added or averaged here; each point represents a number for a state in a given year."


test_wht <- tripTick_abund |> 
    filter(Species == "white",
           State %in% c("GA", "SC"))

ggplot(test_wht) +
    geom_boxplot(aes(x = Wave,
                     y = annland/1000000),
                 fill = NA,
                 outliers = FALSE) +
    geom_jitter(aes(x = Wave,
                     y = annland/1000000,
                   col = Wave),
                size = 2,
                alpha = 0.6) +
    theme_bw() +
    theme(legend.position = "none") +
    labs(y = "Landings (millions of pounds)")
```




## About

**Input `.qmd` file** for this section was: **`r xfun::with_ext(knitr::current_input(), "qmd")`** and can be found in the main directory.  

**Input data file(s)** - in `data/intermediate/adultCommercial_dfs.RData` (the single `.RData` file contains multiple data frames):

-   `landings_annual`
-   `landings_seas`
-   `tripTick_abund`


**Modifications to data:**



**Generated data file(s):**  - fully processed files are in the `data/processed` directory:  

-   `adult.landings.brown_byStateAndShrimpYear.csv`, in the subfolder `moreDetailed`, shows the separate totals by year.  
-   `adult.landings.brown_byShrimpYear.csv` - annual landings were totalled across the two states, to be consistent with how white shrimp landings were reported in our input file.  
-   `adult.landings.white_byShrimpYear.csv` - has both spawning and nonspawning in the same file, as separate `life_stage`s.  


**Other notes about data:**

- none