# Environmental Data  

```{r}
library(tidyverse)
library(patchwork)

source("functions.R")

in_path <- here::here("data", "raw", "environmental")
```


## Winter Temperatures at Charleston Harbor  

winter = dec, jan, feb. 

for brown shrimp, this is all one 'shrimp year'.
white shrimp go through 2 winters: postlarval winter and subadult winter.
subadult winter is the one managers focus on but I will calculate metrics for both.  

See @sec-shrimpyearexpl for more on how these fit into the 'shrimp year' definitions used in this document.  

for shrimp year 1993, postlarval winter is calendar year 1993 and 
subadult winter is calendar year 1994

want min winter temp, mean winter temp, and # days temp < 11C

nursery periods, per Batchelder et al. 2024, are Apr-Jul for brown shrimp
and Jun-Oct for white shrimp

nursery period salinity was important to phenology in that pub
so I'm now pulling out some nursery season info, 2/26/25
based on our definitions, for nursery periods, shrimp year is current year for 
both species  

```{r}
ch <- readxl::read_xlsx(here::here(in_path, "SCDNR_CHS_TempData_ 1950_2024.xlsx"))
# cut out summary rows and column
ch <- ch[1:366, -78]
# pivot
ch2 <- ch |> 
    pivot_longer(-c(Day, Month),
                 names_to = "Year",
                 values_to = "waterTempC") |> 
    mutate(across(everything(), as.numeric))
```

### Consecutive days below thresholds  

There is some debate over which temperature threshold is most important. Here, we compare to each temperature from 7-16C.  

```{r}
ch_winter <- ch2 |> 
    # filter to winter, assign shrimp year, grab only complete shrimp years
    filter(Month %in% c(12, 1, 2)) |> 
    mutate(shrimp_year = case_when(Month == 12 ~ Year + 1,
                                   .default = Year)) |> 
    filter(shrimp_year >= 1951,
           shrimp_year <= 2023) |> 
    group_by(shrimp_year) 

thresholds <- 7:16
consec_out <- list()
for(i in seq_along(thresholds)){
    consec_out[[i]] <- ch_winter |> 
        group_by(shrimp_year) |> 
        summarize(stats = consec_stats(waterTempC, thresholds[i])) |> 
        unnest_wider(stats)
}

tmp_thresh <- bind_rows(consec_out)

tmp_thresh <- tmp_thresh |> 
    rename(calendar_year = shrimp_year)
```




```{r}
ggplot(tmp_thresh,
       aes(x = calendar_year,
           col = threshold)) +
    geom_line(aes(y = totalDays,
                  linetype = "total # days")) +
    geom_line(aes(y = longestSpell,
                  linetype = "longest consecutive # days")) +
    facet_wrap(~threshold, ncol = 3) +
    theme_bw() +
    labs(title = "Winter temperatures at CHS",
         subtitle = "By temperature threshold (panels)",
         x = "Year",
         y = "# days") +
    scale_color_continuous(guide = "none") +
    theme(legend.position = "top")
```

### Temperature Distribution  

```{r}
tmp_distn <- ch2 |> 
    filter(Month %in% c(12, 1, 2)) |> 
    mutate(shrimp_year = case_when(Month == 12 ~ Year + 1,
                                   .default = Year)) |> 
    filter(shrimp_year >= 1951,
           shrimp_year <= 2023) |> 
    summarize(.by = shrimp_year,
              winter_wTemp.CHS_min = min(waterTempC, na.rm = TRUE),
              winter_wTemp.CHS_mean = mean(waterTempC, na.rm = TRUE),
              winter_wTemp.CHS_median = median(waterTempC, na.rm = TRUE),
              winter_wTemp.CHS_max = max(waterTempC, na.rm = TRUE),
              winter_wTemp.CHS_p25 = quantile(waterTempC, probs = 0.25, na.rm = TRUE),
              winter_wTemp.CHS_p75 = quantile(waterTempC, probs = 0.75, na.rm = TRUE),
              winter_wTemp.CHS_sd = sd(waterTempC, na.rm = TRUE),
              winter_wTemp.CHS_iqr = IQR(waterTempC, na.rm = TRUE)) |> 
    rename(calendar_year = shrimp_year)
```

### Apply to different species  

```{r}
to_add <- expand.grid(shrimp_year = 1951:1978,
                      species = c("brown", "white")) |> 
    arrange(shrimp_year, species)

# widen out the threshold data frame to get one row per shrimp year
# first pivot longer so can pair up the temperature with the description
tmp_thresh_wide <- tmp_thresh |> 
    mutate(threshold = paste0("lessThan", threshold)) |> 
    pivot_longer(totalDays:longestSpell,
                 names_to = "stat",
                 values_to = "value") |> 
    pivot_wider(names_from = c(stat, threshold),
                names_sep = "_",
                values_from = value)

# add columns for species and life stage
tmp_thresh_wide <- tmp_thresh_wide |> 
    mutate(shrimpyear.brown_winter = calendar_year,
           shrimpyear.white_postlarvalWinter = calendar_year,
           shrimpyear.white_subadultWinter = calendar_year - 1) |> 
    relocate(all_of(starts_with("shrimpyear")), .after = calendar_year)

# thresholds: split into brown and white ----
# brown
tmp_thresh_brn_wide <- tmp_thresh_wide |> 
    select(shrimp_year = shrimpyear.brown_winter,
           everything(),
           -shrimpyear.white_postlarvalWinter,
           -shrimpyear.white_subadultWinter)
# white postlarval
tmp_thresh_wht_postlarv <- tmp_thresh_wide |> 
    select(shrimp_year = shrimpyear.white_postlarvalWinter,
           everything(),
           -shrimpyear.brown_winter,
           -shrimpyear.white_subadultWinter)
names(tmp_thresh_wht_postlarv)[2:32] <- paste0("postlarvWinter.", names(tmp_thresh_wht_postlarv)[2:32])

# white subadult
tmp_thresh_wht_subad <- tmp_thresh_wide |> 
    select(shrimp_year = shrimpyear.white_subadultWinter,
           everything(),
           -shrimpyear.brown_winter,
           -shrimpyear.white_postlarvalWinter)
names(tmp_thresh_wht_subad)[2:32] <- paste0("subadultWinter.", names(tmp_thresh_wht_subad)[2:32])

# re-combine white
tmp_thresh_wht_all <- full_join(tmp_thresh_wht_postlarv,
                                tmp_thresh_wht_subad,
                                by = "shrimp_year")


# distns: split into brown and white ----
# add columns for species and life stage
tmp_distn <- tmp_distn |> 
    mutate(shrimpyear.brown_winter = calendar_year,
           shrimpyear.white_postlarvalWinter = calendar_year,
           shrimpyear.white_subadultWinter = calendar_year - 1) |> 
    relocate(all_of(starts_with("shrimpyear")), .after = calendar_year)

# brown
tmp_distn_brn <- tmp_distn |> 
    select(shrimp_year = shrimpyear.brown_winter,
           everything(),
           -shrimpyear.white_postlarvalWinter,
           -shrimpyear.white_subadultWinter)
# white postlarval
tmp_distn_wht_postlarv <- tmp_distn |> 
    select(shrimp_year = shrimpyear.white_postlarvalWinter,
           postlarvWinter_calendar_year = calendar_year,
           everything(),
           -shrimpyear.brown_winter,
           -shrimpyear.white_subadultWinter)
names(tmp_distn_wht_postlarv) <- stringr::str_replace(names(tmp_distn_wht_postlarv), "winter", "postlarvWinter")

# white subadult
tmp_distn_wht_subad <- tmp_distn |> 
    select(shrimp_year = shrimpyear.white_subadultWinter,
           subadultWinter_calendar_year = calendar_year,
           everything(),
           -shrimpyear.brown_winter,
           -shrimpyear.white_postlarvalWinter)
names(tmp_distn_wht_subad) <- stringr::str_replace(names(tmp_distn_wht_subad), "winter", "subadultWinter")

# re-combine white
tmp_distn_wht_all <- full_join(tmp_distn_wht_postlarv,
                                tmp_distn_wht_subad,
                                by = "shrimp_year")
```


## Salinity  

Nursery period salinity was important to phenology in Batchelder et al. 2024. We define nursery periods as per that manuscript: Apr-Jul for brown shrimp
and Jun-Oct for white shrimp. I use both 'calendar_year' and 'shrimp_year' here because even though they are the same by our definitions (see @tbl-shrimpYearAssignment), it can be a little difficult to keep track of - so I'm being as explicit as possible.    

Here we pull data from the stations CH07, CH09, CH10, and CH18.  

```{r}
trawl_dat <- read.csv(here::here(in_path,
                                 "SCDNR_EstuarineTrawlShrimpSize.csv"))
```

Salinity at each station was averaged to the day and month. There was at least one anomalously high value, so a line of code removed any values above 50.  

```{r}
# there's a row for every shrimp; condense to one row per trawl  
dat <- trawl_dat |> 
    select(Coll, Year, Month, Day, StationCode, SalinityB) |> 
    filter(StationCode %in% c("CH07", "CH09", "CH10", "CH18")) |> 
    distinct() |> 
    arrange(Year, Month, StationCode)

# for each station, average to the day and to the month
# there was an anomalously high value in here so first
# replace anything over 50 with NA
dat_monthly <- dat |> 
    mutate(SalinityB = case_when(SalinityB > 50 ~ NA_real_,
                                 .default = SalinityB)) |> 
    summarize(.by = c(StationCode, Year, Month, Day),
              sal = mean(SalinityB, na.rm = TRUE)) |> 
    summarize(.by = c(StationCode, Year, Month),
              sal = mean(sal, na.rm = TRUE)) |> 
    pivot_wider(names_from = StationCode,
                names_prefix = "sal_",
                values_from = sal) |> 
    rowwise() |> 
    mutate(sal_mean07.09.10 = mean(c(sal_CH07, sal_CH09, sal_CH10), na.rm = TRUE)) |> 
    ungroup() |>   # leaving out CH18 because it doesn't start until the mid-80s and really affects the mean
    pivot_longer(all_of(starts_with("sal")),
                 names_to = "station",
                 names_prefix = "sal_",
                 values_to = "sal")

# each species has different nursery periods - subset and assign appropriate shrimp years
brown_nursery <- dat_monthly |> 
    filter(Month %in% c(4:7))

brown_nursery_byYear <- brown_nursery |> 
    summarize(.by = c(station, Year),
              sal = mean(sal, na.rm = TRUE))

brown_byYear_wide <- brown_nursery_byYear |> 
    pivot_wider(names_from = station,
                values_from = sal) |> 
    rename(calendar_year = Year) |> 
    mutate(species = "brown",
           shrimp_year = calendar_year) |> 
    relocate(species, calendar_year, shrimp_year)
```

### Brown Shrimp Nursery Period  

```{r}
#| label: fig-brownNurserySalinityAll
#| fig-cap: "All brown shrimp nursery period salinity values by shrimp year"
ggplot(brown_nursery,
       aes(x = Year,
           y = sal,
           col = station,
           shape = station)) +
    geom_point(size = 2) +
    theme_bw() +
    labs(y = "Salinity (ppt)")
```

```{r}
#| label: fig-brownNurserySalinityMean
#| fig-cap: "Mean brown shrimp nursery period salinity values by shrimp year"

ggplot(brown_nursery_byYear,
       aes(x = Year,
           y = sal,
           col = station,
           shape = station)) +
    geom_point(size = 2) +
    theme_bw() +
    labs(y = "Mean Nursery Period Salinity (ppt)")
```


**Correlations between stations**  

Mean winter salinity for each year - we can see that mean nursery period salinity is generally well correlated between stations.   

```{r}
#| label: fig-brownNurserySalinityPairs
#| fig-cap: "Pairs plot of brown shrimp nursery period salinities (mean for each year) at the different stations"

GGally::ggpairs(brown_byYear_wide,
                columns = 2:6)
```


### White Shrimp Nursery Period  

```{r}
white_nursery <- dat_monthly |> 
    filter(Month %in% c(6:10))
white_nursery_byYear <- white_nursery |> 
    summarize(.by = c(station, Year),
              sal = mean(sal, na.rm = TRUE))
white_byYear_wide <- white_nursery_byYear |> 
    pivot_wider(names_from = station,
                values_from = sal) |> 
    rename(calendar_year = Year) |> 
    mutate(species = "white",
           shrimp_year = calendar_year) |> 
    relocate(species, calendar_year, shrimp_year)
```

```{r}
#| label: fig-whiteNurserySalinityAll
#| fig-cap: "All white shrimp nursery period salinity values by shrimp year"


ggplot(white_nursery,
       aes(x = Year,
           y = sal,
           col = station,
           shape = station)) +
    geom_point(size = 2) +
    theme_bw() +
    labs(y = "Salinity (ppt)")

```

```{r}
#| label: fig-whiteNurserySalinityMean
#| fig-cap: "Mean white shrimp nursery period salinity values by shrimp year"

ggplot(white_nursery_byYear,
       aes(x = Year,
           y = sal,
           col = station,
           shape = station)) +
    geom_point(size = 2) +
    theme_bw() +
    labs(y = "Mean Nursery Period Salinity (ppt)")

```

```{r}
#| label: fig-whiteNurserySalinityPairs
#| fig-cap: "Pairs plot of white shrimp nursery period salinities (mean for each year) at the different stations"

GGally::ggpairs(white_byYear_wide,
                columns = 2:6)
```


## Combine all environmental data  

```{r}
# combine salinities
sal_means <- bind_rows(brown_byYear_wide, white_byYear_wide) |> 
    relocate(calendar_year, shrimp_year) |> 
    arrange(calendar_year, species)
names(sal_means)[4:8] <- paste0("nursery_sal.", names(sal_means)[4:8])

# complete the year-species grid
to_add2 <- to_add |> 
    rename(calendar_year = shrimp_year)
sal_means <- bind_rows(to_add2, sal_means)
```

Write out the intermediate data frames - yearly summaries.  

```{r}
save(tmp_thresh, tmp_distn, sal_means,
     file = here::here("data", "intermediate",
                       "environmental_dfs.RData"))
```


```{r}
# split out salinity means 
sal_brn <- sal_means |> 
    filter(species == "brown") |> 
    select(-species)

sal_wht <- sal_means |> 
    filter(species == "white") |> 
    select(-species)


# join it all together 
# brown
env_brn <- full_join(sal_brn, tmp_distn_brn) |> 
    full_join(tmp_thresh_brn_wide) |> 
    arrange(shrimp_year) |> 
    relocate(calendar_year, shrimp_year)

# white
env_wht <- full_join(sal_wht, tmp_distn_wht_all) |> 
    full_join(tmp_thresh_wht_all) |> 
    arrange(shrimp_year) |> 
    select(-calendar_year)

# write out
out_path <- here::here("data", "processed", "environmental")

write.csv(env_brn, 
          here::here(out_path, "brn_environmental.csv"),
          na = "",
          row.names = FALSE)

write.csv(env_wht, 
          here::here(out_path, "wht_environmental.csv"),
          na = "",
          row.names = FALSE)
```



## About

**Input `.qmd` file** for this section was:  **`r xfun::with_ext(knitr::current_input(), "qmd")`**  and can be found in the main directory.  

**Input data file(s)** - as-used; can be found in `data/raw/environmental`:

-   `SCDNR_CHS_TempData_ 1950_2024.xlsx` for temperature  
-   `SCDNR_EstuarineTrawlShrimpSize.csv` for salinity from trawls  


**Modifications to data:**

Summary calculations, described below for each output data frame.    


**Generated data file(s):**  

Data frames were modified as described above, and written to `data/intermediate/environmental_dfs.RData` (the single `.RData` file contains multiple data frames):

-   `tmp_thresh` - counts of days below temperature thresholds for each year. 
   -   `calendar year` represents the year containing January and February. Because we defined winter as December-February, calendar_year 2004 summarizes the time period December 2003 - February 2004.   
   -   `threshold` column contains each integer from 7-16, representing degrees C. 
   -   `totalDays` is the total number of days in that time period below the threshold in the `threshold` column.  
   -   `totalSpells` represents how many times there were at least 7 consecutive days below the given threshold.  
   -   `longestSpell` is the longest stretch of consecutive days (with a minimum of at least 7) within the time period.  
-   `tmp_distn` follows the same `calendar_year` scheme as `tmp_thresh`. In this data frame, there is only one column for each 'calendar_year'. Columns represent which 'shrimp year' the calendar year would fall into for each species/life stage, followed by summary statistics for water temperature over Dec-Feb: min, median, max, mean, sd, iqr, etc.  
-   `sal_means` contains one row for each 'shrimp year' (matches calendar year) and species. Columns are average nursery period salinity from each of 4 stations, and one column containing the mean from 3 of the four stations (one station was less like the others). Nursery periods were defined as in Batchelder et al. 2024; see above for more.    


The data files to be used further in the workflow combined temperature and salinity information for each year x species, and were separated by species. These are both saved in `data/processed/environmental`. Each has a single row per shrimp year. For white shrimp, there are columns for each of 'postlarval winter' and 'subadult winter', along with columns identifying which calendar year was used for the calculations, because the white shrimp life cycle is much trickier than brown:  

-   `brn_environmental.csv`  
-   `wht_environmental.csv`  


**Other notes about data:**
