# Postlarval Stage  

#### Setup  

```{r}
library(dplyr)
library(tidyr)

source(here::here("functions.R"))

in_path <- here::here("data", "intermediate")
out_path <- here::here("data", "processed", "byLifeStage")
out_path_detailed <- here::here(out_path, "moreDetailed")


load(here::here(in_path, "postlarval_dfs.RData"))

postlarv_abund <- postlarv_abund |> 
    select(sample, replicate, date, 
           year, month, day,
           total.ppl = total_ppl_density,
           brown = brown_density,
           white = white_density,
           pink = pink_density) |> 
    mutate(date = lubridate::mdy(date),
           across(c(total.ppl:pink),
                  as.numeric)) |> 
    pivot_longer(c(brown, white, pink),
                 names_to = "species",
                 values_to = "density")
```

#### Compiling to monthly  

```{r}
# compile to monthly

# first average replicates for each sample
postlarv_abund_monthly <- postlarv_abund |> 
    summarize(.by = c(year, month, day, sample, species),
              density = mean(density, na.rm = TRUE)) |> 
    # now go up to month
    summarize(.by = c(year, month, species),
              density = mean(density, na.rm = TRUE)) |> 
    arrange(year, month, species) |> 
    mutate(survey = "NInlet",
           life_stage = "postlarval",
           abundance_measure = "density",
           abundance = density) |> 
    filter(year != 1900,
           species %in% c("brown", "white"))

postlarv_abund_monthly <- add_cohorts(postlarv_abund_monthly, "postlarvae")

postlarv_abund_monthly <- postlarv_abund_monthly |> 
    select(all_of(sel_order_monthly))
```

#### Compiling to shrimp year  

```{r}
# annual file - by shrimp year
# removing late postlarvae here - shrimp_year of 9999
postlarv_yearly <- postlarv_abund_monthly |> 
    filter(!is.na(shrimp_year),
           shrimp_year != 9999) |> 
    summarize(.by = c(survey, species, life_stage, shrimp_year, abundance_measure),
              abundance = sum(abundance, na.rm = TRUE)) |> 
    mutate(sqrt_abundance = sqrt(abundance)) |> 
    select(all_of(sel_order_annual)) |> 
    group_by(species) |> 
    mutate(abundanceIndex_fullTimePeriod = as.vector(scale(sqrt_abundance))) |> 
    ungroup()

abund_commontime <- postlarv_yearly |> 
    select(species, shrimp_year, sqrt_abundance) |> 
    filter(shrimp_year %in% common_shrimpYears) |> 
    group_by(species) |> 
    mutate(abundanceIndex_commonTimePeriod = as.vector(scale(sqrt_abundance))) |> 
    ungroup()

postlarv_yearly <- left_join(postlarv_yearly,
                             abund_commontime)
```

#### Saving files out  

```{r}
# write out

# monthly
write.csv(postlarv_abund_monthly,
          here::here(out_path_detailed, "postlarval_byMonth.csv"))
# yearly
write.csv(postlarv_yearly,
          here::here(out_path, "postlarval_byShrimpYear.csv"))
```


## Tabular summaries of output datasets  

### Monthly  

<details>

::: {.content-visible when-format="html"}
<summary>Click to expand</summary>
:::

```{r}
#| label: tbl-monthlyAbund
#| tbl-cap: "Summary of monthly postlarval abundance data"
skimr::skim_without_charts(postlarv_abund_monthly)
```

</details>

### Yearly  

<details>

::: {.content-visible when-format="html"}
<summary>Click to expand</summary>
:::

```{r}
#| label: tbl-yearlyAbund
#| tbl-cap: "Summary of postlarval abundance data by shrimp year"
skimr::skim_without_charts(postlarv_yearly)
```

</details>

## About

**Input `.qmd` file** for this section was: **`r xfun::with_ext(knitr::current_input(), "qmd")`** and can be found in the main directory.  

**Input data file(s)** - in `data/intermediate/postlarval_dfs.RData` (the single `.RData` file contains multiple data frames):

-   `postlarv_abund`
-   `postlarv_size`  (this data frame was ignored)  


**Modifications to data:**

-  there were some 'year' entries of 1900. These were removed. No data or dates were associated with these rows; presumably they were extra rows in the spreadsheet to aid in data entry.   
-  removed data for pink shrimp.  


**Generated data file(s):**  - fully processed files are in the `data/processed` directory:  

-  `postlarval_byMonth.csv`, in the subfolder `moreDetailed`, is summarized to species and month within year.  
-  `postlarval_byShrimpYear.csv` is summarized to species and shrimp year.  

**Other notes about data:**

- there were some 'year' entries of 1900. These were removed. No data or dates were associated with these rows; presumably they were extra rows in the spreadsheet to aid in data entry.   
