# Juvenile  

There are two datasets that represent juveniles: SCDNR's Creek Trawls, and NIW NERR's Oyster Landing seines. 

Setup 

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

source(here::here("functions.R"))

in_path <- here::here("data", "intermediate")
out_path <- here::here("data", "processed")

load(here::here(in_path, "juvenile_dfs.RData"))
```


#### SC Creek Trawls - compile to monthly  

```{r}
names(juv_sc_cpue) <- tolower(names(juv_sc_cpue))
juv_sc_cpue <- filter(juv_sc_cpue,
                      species %in% c("brown", "white"))

# SC Creek Trawls
# compile to monthly

# no replicates to average; just go straight to monthly avg
juv_cpue_monthly <- juv_sc_cpue |> 
    summarize(.by = c(year, month, species),
              cpue = mean(cpue, na.rm = TRUE)) |> 
    mutate(survey = "SC Creek Trawls",
           life_stage = "juvenile",
           abundance_measure = "CPUE",
           abundance = cpue)

juv_cpue_monthly <- add_cohorts(juv_cpue_monthly, "juvenile") |> 
    select(all_of(sel_order_monthly))
```

#### Oyster Landing - compile to monthly  

```{r}
# Oyster Landing
# compile to monthly

names(juv_ol_count) <- tolower(names(juv_ol_count))

# no replicates to deal with

juv_ol_monthly <- juv_ol_count |> 
    summarize(.by = c(year, month, species),
              count = mean(count, na.rm = TRUE)) |> 
    mutate(survey = "OysterLanding",
           life_stage = "juvenile",
           abundance_measure = "count",
           abundance = count)

juv_ol_monthly <- add_cohorts(juv_ol_monthly, "juvenile") |> 
    arrange(year, month, species) |> 
    select(all_of(sel_order_monthly))
```

#### Combine monthly data frames  

```{r}
juv_monthly_combined <- bind_rows(juv_cpue_monthly,
                                  juv_ol_monthly) |> 
    arrange(year, month, species, survey)
```

Make sure numbers are generally on the same scale before averaging.  

```{r}
#| label: fig-juvCombined
#| fig-cap: "Jittered graphic of the square-root of abundance by species and survey. The different surveys used different measures of abundance (SC Creek Trawls used CPUE; Oyster Landing used count) but they appear to be on the same scale, so averaging them should be fine."
ggplot(juv_monthly_combined,
       aes(x = species, y = sqrt(abundance), col = survey)) +
    geom_jitter()
```

## Compile to Shrimp Year  

First, get annual totals by survey.   


```{r}
# annual file - by shrimp year
# removing late postlarvae here - shrimp_year of 9999
juv_combined_byShrimpYear <- juv_monthly_combined |> 
    group_by(life_stage, survey, species, shrimp_year, abundance_measure) |> 
    summarize(abundance = sum(abundance, na.rm = TRUE)) |> 
    ungroup() |> 
    mutate(sqrt_abundance = sqrt(abundance)) |> 
    select(all_of(sel_order_annual)) |> 
    arrange(shrimp_year, species, survey)
```

Now average the annual values across surveys.

```{r}
juv_combined_byShrimpYear_averaged <- juv_combined_byShrimpYear |> 
    group_by(species, life_stage, shrimp_year) |> 
    summarize(abundance = mean(abundance, na.rm = TRUE)) |> 
    ungroup() |> 
    mutate(sqrt_abundance = sqrt(abundance),
           survey = "SC Creek Trawls and OL Seines",
           abundance_measure = "mean across surveys; count matched CPUE in magnitude") |> 
    select(all_of(sel_order_annual))
```

#### Write out.  

```{r}
# write out

# monthly
write.csv(juv_monthly_combined,
          here::here(out_path, "juvenile_bySurveyAndMonth.csv"))
# yearly
write.csv(juv_combined_byShrimpYear,
          here::here(out_path, "juvenile_bySurveyAndShrimpYear.csv"))
write.csv(juv_combined_byShrimpYear_averaged,
          here::here(out_path, "juvenile_byShrimpYear.csv"))
```


## Tabular summaries of output datasets  

### Monthly  

<details>

::: {.content-visible when-format="html"}
<summary>Click to expand</summary>
:::

```{r}
#| label: tbl-monthlyJuvSC
#| tbl-cap: "Summary of monthly juvenile abundance data, SC Creek Trawls"
skimr::skim_without_charts(juv_cpue_monthly)
```

```{r}
#| label: tbl-monthlyJuvOL
#| tbl-cap: "Summary of monthly juvenile abundance data, Oyster Landing"
skimr::skim_without_charts(juv_ol_monthly)
```

</details>

### Yearly  

<details>

::: {.content-visible when-format="html"}
<summary>Click to expand</summary>
:::

```{r}
#| label: tbl-yearlyJuv
#| tbl-cap: "Summary of juvenile abundance data by shrimp year, after averaging SC and OL surveys."
skimr::skim_without_charts(juv_combined_byShrimpYear_averaged)
```

</details>

## About

**Input `.qmd` file** for this section was: **`r xfun::with_ext(knitr::current_input(), "qmd")`** and can be found in the main directory.  

**Input data file(s)** - in `data/intermediate/juvenile_dfs.RData` (the single `.RData` file contains multiple data frames):

-   `juv_sc_cpue`
-   `juv_sc_size`
-   `juv_ol_count`
-   `juv_ol_size`


**Modifications to data:**



**Generated data file(s):**  - fully processed files are in the `data/processed` directory:  

-  `juvenile_bySurveyAndMonth.csv` is summarized to species and month within year, by survey.  
-  `juvenile_bySurveyAndShrimpYear.csv` is summarized to species and shrimp year, by survey. 
-  `juvenile_byShrimpYear.csv` is summarized to species and shrimp year, as the average shrimp-year abundance across the two surveys. This is the file that will be used in further analyses.  

**Other notes about data:**

- none