# Adults - landings and trip tickets  

#### Setup 

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

source(here::here("functions.R"))

in_path <- here::here("data", "intermediate")
out_path <- here::here("data", "processed", "byLifeStage")
out_path_detailed <- here::here(out_path, "moreDetailed")

load(here::here(in_path, "adultCommercial_dfs.RData"))
```

## Landings Data  

### Brown Shrimp  

Use annual data for brown shrimp, because it corresponds to the life cycle and shrimp year.  Three states are represented in this dataset. NC is not represented elsewhere, so it is removed and we will focus only on GA and SC. 

```{r}
brn_landings <- landings_annual |> 
    janitor::clean_names() |> 
    filter(species == "brown",
           state != "NORTH CAROLINA") |> 
    select(shrimp_year = year,
           species,
           state,
           pounds)

brn_abund <- brn_landings |> 
    mutate(survey = case_when(state == "GEORGIA" ~ "GA",
                              state == "SOUTH CAROLINA" ~ "SC",
                              .default = "PROBLEM"),
           life_stage = "adult.landings",
           abundance_measure = "pounds",
           abundance = pounds,
           sqrt_abundance = sqrt(abundance)) |> 
    select(all_of(sel_order_annual)) |> 
    arrange(shrimp_year, species, survey)
```

```{r}
#| label: fig-brnAbundByState
#| fig-caption: "Brown shrimp annual landings by state"

ggplot(brn_abund,
       aes(x = shrimp_year,
           y = abundance,
           col = survey)) +
    geom_point(size = 2)
```

Total over the two states (because the data we have for white shrimp is totals across the states, not averages) and generate abundance index.  

```{r}
brn_byShrimpYear <- brn_abund |> 
    group_by(species, life_stage, shrimp_year) |> 
    summarize(abundance = sum(abundance, na.rm = TRUE)) |> 
    ungroup() |> 
    mutate(sqrt_abundance = sqrt(abundance),
           survey = "Landings, GA and SC",
           abundance_measure = "total over 2 states") |> 
    select(all_of(sel_order_annual)) |> 
    group_by(species) |> 
    mutate(abundanceIndex_fullTimePeriod = as.vector(scale(sqrt_abundance))) |> 
    ungroup()

brn_abund_commontime <- brn_byShrimpYear |> 
    select(species, shrimp_year, sqrt_abundance) |> 
    filter(shrimp_year %in% common_shrimpYears) |> 
    group_by(species) |> 
    mutate(abundanceIndex_commonTimePeriod = as.vector(scale(sqrt_abundance))) |> 
    ungroup()

brn_byShrimpYear <- left_join(brn_byShrimpYear, 
                              brn_abund_commontime)
```

Write out.  

```{r}
# by state
write.csv(brn_abund,
          here::here(out_path_detailed,
                     "adult.landings.brown_byStateAndShrimpYear.csv"),
          na = "",
          row.names = FALSE)

# combined with index
write.csv(brn_byShrimpYear,
          here::here(out_path,
                     "adult.landings.brown_byShrimpYear.csv"),
          na = "",
          row.names = FALSE)
```


### White Shrimp  

Using seasonal data for white shrimp, to separate non-spawning adults (time period 2 for a calendar year, season Fall, Months Aug-Dec) from spawners (time period 1 of the next calendar year, season Winter-Spring-Summer, months Jan-July).   

The input file contained heads-off landings from SC and GA.  

```{r}
# Fall = Aug-Dec  - make shrimp_year =  year
# Win_Spr_Sum = Jan-Jul - make shrimp_year = year - 1

wht_landings <- landings_seas |> 
    mutate(shrimp_year = case_when(season == "Fall" ~ Year,
                                   .default = Year - 1),
           life_stage = case_when(season == "Fall" ~ "adult.nonspawning.landings",
                                  .default = "adult.spawning.landings"),
           survey = "landings",
           species = "white",
           abundance_measure = "head-off weight, sum of SC and GA",
           abundance = lbs_headOff,
           sqrt_abundance = sqrt(abundance)) |> 
    filter(shrimp_year >= 1979,
           shrimp_year <= 2022) |> 
    janitor::clean_names() |> 
    select(any_of(sel_order_monthly),
           season, sqrt_abundance) |> 
    relocate(season, .after = "year")
```

Generate abundance index. Handling a little differently because there are two life stages in here (spawning and non-spawning adults).  

```{r}
# abundance index, full time
wht_abund <- wht_landings |> 
    group_by(species, life_stage) |> 
    mutate(abundanceIndex_fullTimePeriod = as.vector(scale(sqrt_abundance))) |> 
    ungroup()

# abundance index, common time
wht_abund_commonTime <- wht_abund |> 
    select(species, life_stage, shrimp_year, sqrt_abundance) |> 
    filter(shrimp_year %in% common_shrimpYears) |> 
    group_by(species, life_stage) |> 
    mutate(abundanceIndex_commonTimePeriod = as.vector(scale(sqrt_abundance))) |> 
    ungroup()

wht_abund <- left_join(wht_abund, 
                       wht_abund_commonTime) |> 
    rename(calendar_year = year) |> 
    select(survey, species, shrimp_year, life_stage,
           calendar_year, season,
           everything())
```

```{r}
#| label: fig-whtAbundIndex
#| fig-cap: "Abundance Index for white shrimp, by spawning vs. non-spawning adults"


ggplot(wht_abund,
       aes(x = shrimp_year,
           y = abundanceIndex_fullTimePeriod,
           col = life_stage)) +
    geom_point(size = 2) +
    geom_line() +
    theme_bw()
```


```{r}
#| eval: false 

ggplot(wht_abund) +
    geom_histogram(aes(x = abundanceIndex_fullTimePeriod)) +
    facet_wrap(~life_stage)
```

Write out white shrimp data. Both 'adult' life stages are in the single file.  

```{r}
write.csv(wht_abund,
          here::here(out_path,
                     "adult.landings.white_byShrimpYear.csv"),
          na = "",
          row.names = FALSE)
```



## Trip Ticket Data  

I intended to compile to annual totals, but in early years there are big differences in reporting frequencies. So the only files I will write out will contain the 'Wave', which is the 2-month period of reporting. I will not compile to annual values or abundance indices.  

Would probably be best to pull out a wave or two and use those as an index. Waves 3, 4, and 5 seem to be pretty consistently reported after about 2000.  

### Brown Shrimp

```{r}
brn_tt_byState <- tripTick_abund |> 
    filter(Species == "brown",
           State %in% c("GA", "SC")) |> 
    mutate(shrimp_year = Year,
           life_stage = "adult.triptickets",
           survey = paste0("Trip tickets, ", State),
           pseudodate = lubridate::ymd(paste(Year, Wave_start, 15))) |> 
    select(pseudodate, Year, Wave, shrimp_year, survey, life_stage, 
           annland, numdeal, numfisher, CPUE_approx) |> 
    arrange(Year, Wave)

# add the states together, within wave
brn_tt <- brn_tt_byState |> 
    summarize(.by = c(pseudodate, shrimp_year, Wave, life_stage),
              annland = sum(annland, na.rm = TRUE),
              numdeal = sum(numdeal, na.rm = TRUE),
              numfisher = sum(numfisher, na.rm = TRUE)) |> 
    mutate(CPUE_approx = annland/numfisher,
           survey = "Trip tickets, GA + SC total")
```

```{r}
#| label: fig-brnCPUEbyWave
#| fig-cap: "Approximate CPUE for GA & SC combined (added together) for brown shrimp by time period. Notice the vastly different reporting frequencies before about 2000."

ggplot(brn_tt,
       aes(x = pseudodate,
           y = CPUE_approx)) +
    geom_line(col = "gray50") +
    geom_point(aes(col = Wave,
                   shape = Wave),
               size = 2) +
    theme_bw() +
    theme(legend.position = "bottom") +
    labs(title = "Brown Shrimp",
         x = "Date")
```


```{r}
# by state & wave
write.csv(brn_tt_byState,
          here::here(out_path_detailed,
                     "adult.triptickets.brown_01byStateAndWave.csv"),
          na = "",
          row.names = FALSE)

# by state & shrimp year
write.csv(brn_tt,
          here::here(out_path_detailed,
                     "adult.triptickets.brown_02byWave.csv"),
          na = "",
          row.names = FALSE)
```


### White Shrimp  

For whites, ideally I would compile to the same seasons reported in landings: Jan-Jul and Aug-Dec. However the waves here split; Jul-Aug are combined. Per advice from a work group member, July landings are negligible and this wave should be dominated by August landings, so we will count them with Fall/non-spawning shrimp.  

There are most likely differences in when reporting happened seasonally, as with brown shrimp, so I will again only compile to "Wave".  

```{r}
wht_tt_byState <- tripTick_abund |> 
    filter(Species == "white",
           State %in% c("GA", "SC")) |> 
    mutate(shrimp_year = case_when(WAVE <= 3 ~ Year - 1,
                                   WAVE >= 4 ~ Year),
           life_stage = case_when(WAVE <= 3 ~ "adult.spawning.triptickets",
                                  WAVE >= 4 ~ "adult.nonspawning.triptickets"),
           survey = paste0("Trip tickets, ", State),
           pseudodate = lubridate::ymd(paste(Year, Wave_start, 15))) |> 
    select(Year, pseudodate, Wave, shrimp_year, survey, life_stage, 
           annland, numdeal, numfisher, CPUE_approx)

# add the states together, within wave
wht_tt <- wht_tt_byState |> 
    summarize(.by = c(pseudodate, shrimp_year, Wave, life_stage),
              annland = sum(annland, na.rm = TRUE),
              numdeal = sum(numdeal, na.rm = TRUE),
              numfisher = sum(numfisher, na.rm = TRUE)) |> 
    mutate(CPUE_approx = annland/numfisher,
           survey = "Trip tickets, GA + SC total")
```

```{r}
#| label: fig-whtCPUEApproxByWave
#| fig-cap: "Approximate CPUE for GA & SC combined (added together) for white shrimp by time period. Notice the vastly different reporting frequencies before about 2000."

ggplot(wht_tt,
       aes(x = pseudodate,
           y = CPUE_approx)) +
    geom_line(col = "gray50") +
    geom_point(aes(col = life_stage,
                   shape = Wave),
               size = 2) +
    theme_bw() +
    theme(legend.position = "bottom") +
    labs(title = "White Shrimp",
         x = "Date")
```

Looks like there was a lot going on before about 1995 so let's filter that data.  

```{r}
#| label: fig-whtCPUEApproxByWave2
#| fig-cap: "Approximate CPUE for GA & SC combined (added together) for white shrimp by time period, starting in 1995"

ggplot(filter(wht_tt, shrimp_year >= 1995),
       aes(x = pseudodate,
           y = CPUE_approx)) +
    geom_line(col = "gray50") +
    geom_point(aes(col = life_stage,
                   shape = Wave),
               size = 2) +
    theme_bw() +
    theme(legend.position = "bottom") +
    labs(title = "White Shrimp")
```

There are several years where Wave 2 is missing (triangle), but 1, 3, 4, 5, and 6 seem pretty consistent in this time period (post-1995).  


**All years are included in these output files**.  

```{r}
# by state & wave
write.csv(wht_tt_byState,
          here::here(out_path_detailed,
                     "adult.triptickets.white_01byStateAndWave.csv"),
          na = "",
          row.names = FALSE)

# by state & shrimp year
write.csv(wht_tt,
          here::here(out_path_detailed,
                     "adult.triptickets.white_02byWave.csv"),
          na = "",
          row.names = FALSE)
```


## About

**Input `.qmd` file** for this section was: **`r xfun::with_ext(knitr::current_input(), "qmd")`** and can be found in the main directory.  

**Input data file(s)** - in `data/intermediate/adultCommercial_dfs.RData` (the single `.RData` file contains multiple data frames):

-   `landings_annual`
-   `landings_seas`
-   `tripTick_abund`


**Modifications to data:**  

-   pink shrimp removed. 
-   any areas other than GA and SC were removed.  
-   `pseudodate` column was created based on the year and first month of each 'Wave' of reporting, in order to make time series effectively.


**Generated data file(s):**  - fully processed files are in the `moreDetailed` folder of the `data/processed` directory. Abundance indices were not generated due to inconsistency in the consituents (which Waves had values through time):  

-   `adult.triptickets.brown_01byStateAndWave.csv` - abundances and number reporting for GA and SC separately by Wave and year.  
-   `adult.triptickets.brown_02byWave.csv` - abundances and reporters from the two states are added together for Wave and year.  
-   `adult.triptickets.white_01byStateAndWave.csv` - same as brown file  
-   `adult.triptickets.white_02byWave.csv` - same as brown file  

**Other notes about data:**

- none