# Adults - fisheries-independent  

This file focuses on SEAMAP data.  

#### Setup 

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

source(here::here("functions.R"))

in_path <- here::here("data", "intermediate")
out_path <- here::here("data", "processed", "byLifeStage")
out_path_detailed <- here::here(out_path, "moreDetailed")

load(here::here(in_path, "adultSeamap_dfs.RData"))
```

```{r}
adults <- adsm_long |> 
    janitor::clean_names() |> 
    filter(species %in% c("brown", "white")) 
```

#### Compiling to season  

SEAMAP does seasonal sampling; not monthly.  


# stopping here for the night  

Need to deal with seasons, get shrimp year assigned, and spawning vs. non-spawning for white shrimp.  

```{r}
# average to Event,
# then average to season by state,
# then calculate annual totals


# ad_sm_seasonal <- ad_sm_long |> 
#     mutate(season = case_when(month %in% c(4, 5) ~ "Spring",
#                               month %in% c(7, 8) ~ "Summer",
#                               month %in% c(9, 10, 11) ~ "Fall",
#                               .default = "JuneProbably")) |> 
#     filter(season != "JuneProbably") |> 
#     # average to event
#     summarize(.by = c(species, year, month, shrimp_year, season, region,
#                       event),
#               count = mean(count, na.rm = TRUE),
#               weight = mean(weight, na.rm = TRUE)) |> 
#     # average to season within state
#     summarize(.by = c(species, year, shrimp_year, season, month, region),
#               count = mean(count, na.rm = TRUE),
#               weight = mean(weight, na.rm = TRUE)) |> 
#     mutate(survey = paste0("SEAMAP - ", region),
#            life_stage = "adult",
#            abundance_measure = "weight",
#            abundance = weight) |> 
#     select(survey, species, life_stage, year, month, season, shrimp_year, abundance_measure, abundance)
# 
# # write that out to the seasonal folder
# write.csv(ad_sm_seasonal,
#           here::here(path_out, "monthly or seasonal",
#                      "04_adultSEAMAPbyRegion_bySeason.csv"),
#           na = "",
#           row.names = FALSE)

```



## About

**Input `.qmd` file** for this section was: **`r xfun::with_ext(knitr::current_input(), "qmd")`** and can be found in the main directory.  

**Input data file(s)** - in `data/intermediate/adultSeamap_dfs.RData` (the single `.RData` file contains multiple data frames):

-   `seamap` - the unfiltered data frame
-   `adsm_wide`
-   `adsm_long`


**Modifications to data:**



**Generated data file(s):**  - fully processed files are in the `data/processed` directory:  


**Other notes about data:**

- none