# Adults - fisheries-independent  

This file focuses on SEAMAP data.  

#### Setup 

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

source(here::here("functions.R"))

in_path <- here::here("data", "intermediate")
out_path <- here::here("data", "processed", "byLifeStage")
out_path_detailed <- here::here(out_path, "moreDetailed")

load(here::here(in_path, "adultSeamap_dfs.RData"))
```

```{r}
# filter the dataset
adults <- adsm_long |> 
    janitor::clean_names() |> 
    filter(species %in% c("brown", "white"),
           region %in% c("GEORGIA", "SOUTH CAROLINA")) 

# add shrimp year
adults <- add_cohorts(adults, "adult") 

# add life stage
adults <- adults |> 
    mutate(life_stage = case_when(species == "brown" ~ "adult",
                                  species == "white" & month <= 5 ~ "adult.spawning",
                                  .default = "adult.nonspawning"),
           region = case_when(region == "GEORGIA" ~ "GA",
                              .default = "SC"))
```

#### Compiling to season  

SEAMAP does seasonal sampling; not monthly.  

So here, I average to sampling event; then average to season within state; and then calculate annual totals.
 

```{r}
ad_seasonal <- adults |>
    mutate(season = case_when(month %in% c(4, 5) ~ "Spring",
                              month %in% c(7, 8) ~ "Summer",
                              month %in% c(9, 10, 11) ~ "Fall",
                              .default = "JuneProbably")) |>
    filter(season != "JuneProbably") |>
    # average to event
    summarize(.by = c(species, life_stage, 
                      year, month, shrimp_year, 
                      season, region, event),
              count = mean(count, na.rm = TRUE),
              weight = mean(weight, na.rm = TRUE)) |>
    # average to season within state
    summarize(.by = c(species, year, life_stage, shrimp_year, season, region),
              count = mean(count, na.rm = TRUE),
              weight = mean(weight, na.rm = TRUE)) |>
    mutate(survey = paste0("SEAMAP - ", region),
           abundance_measure = "weight",
           abundance = weight) |>
    select(survey, species, life_stage, year, season, shrimp_year, abundance_measure, abundance)  # not using 'sel_order_monthly' because we have season, not month

```

Yearly totals by state.  

```{r}
ad_yearly_state <- ad_seasonal |> 
    summarize(.by = c(survey, species, life_stage, shrimp_year, abundance_measure),
              abundance = sum(abundance, na.rm = TRUE)) 
```

```{r}
#| label: fig-seamapAnnualByLifeStage
#| fig-cap: "Time series of total abundance (weight) by species x life stage, colored by state, in SEAMAP surveys."

ad_yearly_state |> 
    mutate(sp.stage = paste(species, life_stage)) |> 
ggplot(aes(x = shrimp_year,
           y = abundance,
           col = survey)) +
    geom_line() +
    geom_point(size = 2) +
    facet_wrap(~sp.stage, scales = "free", ncol = 1) +
    theme_bw() +
    theme(legend.position = "bottom")
```

Average the states each year and calculate abundance index.  

```{r}
ad_yearly_final <- ad_yearly_state |> 
    summarize(.by = c(species, life_stage, abundance_measure, shrimp_year),
              abundance = mean(abundance, na.rm = TRUE)) |>
    group_by(species, life_stage, abundance_measure) |> 
    mutate(sqrt_abundance = sqrt(abundance),
           survey = "SEAMAP - GA and SC",
           abundanceIndex_fullTimePeriod = as.vector(scale(sqrt_abundance))) |> 
    ungroup() |> 
    select(all_of(sel_order_annual),
           everything())

abund_commontime <- ad_yearly_final |> 
    select(species, life_stage, shrimp_year, sqrt_abundance) |> 
    filter(shrimp_year %in% common_shrimpYears) |> 
    group_by(species) |> 
    mutate(abundanceIndex_commonTimePeriod = as.vector(scale(sqrt_abundance))) |> 
    ungroup()

ad_yearly_final <- left_join(ad_yearly_final,
                             abund_commontime)
```


Write out the files.  

```{r}
# by state and season
write.csv(ad_seasonal,
          here::here(out_path_detailed, 
                     "adult.seamap_byStateAndSeason.csv"),
          na = "",
          row.names = FALSE)

# shrimp year total by state
write.csv(ad_yearly_state,
          here::here(out_path_detailed,
                     "adult.seamap_byStateAndShrimpYear.csv"),
          na = "",
          row.names = FALSE)

# with abundance index
write.csv(ad_yearly_final,
          here::here(out_path,
                     "adult_byShrimpYear.csv"),
          na = "",
          row.names = FALSE)
```


## About  

**Input `.qmd` file** for this section was: **`r xfun::with_ext(knitr::current_input(), "qmd")`** and can be found in the main directory.  

**Input data file(s)** - in `data/intermediate/adultSeamap_dfs.RData` (the single `.RData` file contains multiple data frames):

-   `seamap` - the unfiltered data frame
-   `adsm_wide`
-   `adsm_long`


**Modifications to data:**

-   filtered to only brown and white shrimp  
-   filtered to only SC and GA  
-   averaged to sampling event, then to season, by state. Summed seasonal averages by shrimp year, by state. Then averaged annual totals and calculated abundance index.  


**Generated data file(s):**  - fully processed files are in the `data/processed` directory:  

-   `adult.seamap_byStateAndSeason.csv` - in `moreDetailed` folder - abundance averaged to season within state, by species and life stage.  
-   `adult.seamap_byStateAndShrimpYear.csv` - in `moreDetailed` folder - contains annual totals by species, life stage, shrimp year, and state.  
-   `adult_byShrimpYear.csv` - contains average of annual totals by shrimp year and calculated abundance index, by species, life stage, and shrimp year.  


**Other notes about data:**

-   did not worry about order of rows in output files, but all life stages are represented equally.  