# Postlarval  

```{r}
library(tidyverse)
library(patchwork)
library(readxl)
library(ggbeeswarm)

source("functions.R")

in_path <- here::here("data", "raw", "postlarv")

surv <- "N. Inlet Postlarval Sampling"
```

```{r}
postlarv_abund <- read.csv(here::here(in_path, "Penaeus_PostLarvae_NInlet_1981_2017_wide_kac.csv"))
postlarv_size <- read_xlsx(here::here(in_path, "Penaeus_Postlarval_Lengths_NInlet_1981 - 2017.xlsx"))
```

##  Dataset  

The only dataset representing postlarvae is from North Inlet-Winyah Bay. Two files are used: `Penaeus_PostLarvae_NInlet_1981_2017_wide_kac.csv` and `Penaeus_Postlarval_Lengths_NInlet_1981 - 2017.xlsx`. The latter file contains sizes, which were explored in this portion of the project but not expanded upon during later compiling stages. 


```{r}
postlarv_abund <- postlarv_abund |> 
    mutate(date2 = lubridate::mdy(date),
           across(c(total_ppl_density:pink_density),
                  as.numeric))

postlarv_size <- postlarv_size |> 
    rename(Species = `Shrimp species`,
           Rostrum_teeth = `Rostrum teeth`,
           Rostrum_length = `Rostrum Length`,
           Telson_length = `Telson length`,
           Uropod_length = `Uropod length`) |> 
    mutate(across(c(Rostrum_teeth:Uropod_length),
                  as.numeric),
           Year = lubridate::year(Date),
           Month = lubridate::month(Date))

postlarv_size_brown <- postlarv_size |> 
    filter(Species == "brown")

postlarv_size_white <- postlarv_size |> 
    filter(Species == "white")

# save out these data frames
save(postlarv_abund, postlarv_size,
     file = here::here("data",
                       "intermediate",
                       "postlarval_dfs.RData"),
     compress = "xz")
```

## Tabular summary of abundance   

```{r}
#| label: tbl-abundance
#| tbl-cap: "Abundance data frame summary"
skimr::skim_without_charts(postlarv_abund)
```

## Tabular summaries of size   

### Brown Shrimp  

```{r}
#| label: tbl-size-brown
#| tbl-cap: "Brown shrimp size summary"
skimr::skim_without_charts(postlarv_size_brown)
```

### White Shrimp  

```{r}
#| label: tbl-size-white
#| tbl-cap: "White shrimp size summary"
skimr::skim_without_charts(postlarv_size_white)
```

## Graphics - Size Distributions  

In these plots, there is a panel for each shrimp species. Size is represented on the y-axis, and a cloud of points called a "beeswarm" has representation for every data point. Each point is colored by the numeric month in which the individual was caught, allowing us to see differences in size at different parts of the year and life-cycle.  

There are three different measures of size in this dataset: Rostrum length, Telson length, and Uropod length. A graph is below for each.  

```{r}
#| label: fig-beeswarm-rostrum
#| fig-cap: "Beeswarm plot of rostrum length in postlarval shrimp. Points represent measurements of individual shrimp, and are colored by the month in which they were captured. The y-axis represents length. More points spread out along the x-axis for a given length means that there were more individuals of that length captured than regions where points remain closer to the center."

beeswarm_sizeDistn(postlarv_size, 
               x = "Species", y = "Rostrum_length",
               month = "Month",
               survey = surv)
```

```{r}
#| label: fig-beeswarm-telson
#| fig-cap: "Beeswarm plot of telson length in postlarval shrimp. Points represent measurements of individual shrimp, and are colored by the month in which they were captured. The y-axis represents length. More points spread out along the x-axis for a given length means that there were more individuals of that length captured than regions where points remain closer to the center."

beeswarm_sizeDistn(postlarv_size, 
               x = "Species", y = "Telson_length",
               month = "Month",
               survey = surv)
```

```{r}
#| label: fig-beeswarm-uropod
#| fig-cap: "Beeswarm plot of uropod length in postlarval shrimp. Points represent measurements of individual shrimp, and are colored by the month in which they were captured. The y-axis represents length. More points spread out along the x-axis for a given length means that there were more individuals of that length captured than regions where points remain closer to the center."

beeswarm_sizeDistn(postlarv_size, 
               x = "Species", y = "Uropod_length",
               month = "Month",
               survey = surv)
```

With brown shrimp, looks like we've got a couple of size classes in here - bigger shrimp earlier in the year, and smaller ones that are mostly later months.

The plot below shows the entire time series, with date (as year-month) along the x-axis and uropod length on the y-axis. Points are again colored by month.  

```{r}
#| label: fig-length-timeseries
#| fig-cap: "Time series plot of uropod length in postlarval shrimp. Points represent measurements of individual shrimp, and are colored by the month in which they were captured."

ggplot(postlarv_size) +
    geom_point(aes(x = Date,
                   y = Uropod_length,
                   col = Month)) +
    facet_wrap(~Species, scales = "free_y", ncol = 1) +
    khroma::scale_color_YlOrBr()
```


## Graphics - Abundance and Size Boxplots  

```{r}
#| label: fig-abund-size
#| fig-cap: "Boxplots representing shrimp size and abundance Note the log10-scaled y-axes."

p1 <- ggplot(postlarv_size) +
    geom_boxplot(aes(x = factor(Month),
                     y = Uropod_length,
                     fill = Month)) +
    facet_wrap(~Species, ncol = 1) +
    khroma::scale_fill_YlOrBr() +
    theme(legend.position = "none") +
    labs(title = "Size by time of year",
         x = "Month")


postlarv_abund_long <- postlarv_abund |> 
    select(Date = date2,
           Year = year,
           Month = month,
           brown = brown_density,
           white = white_density,
           pink = pink_density) |> 
    pivot_longer(c(brown, white, pink),
                 names_to = "Species",
                 values_to = "Density") |> 
    mutate(Density2 = Density + 0.001)

p2 <- ggplot(postlarv_abund_long) +
    geom_boxplot(aes(x = factor(Month),
                     y = Density2,
                     fill = Month)) +
    facet_wrap(~Species, ncol = 1) +
    scale_y_log10() +
    khroma::scale_fill_YlOrBr() +
    theme(legend.position = "none") +
    labs(title = "Abundance by time of year",
         x = "Month")

p1 + p2
```

So even though we're seeing that pattern in brown shrimp of being so much bigger early in the year and smaller later, they're just really not all that common later in the year.  
