---
format:
    html:
        fig-width: 7
        fig-height: 7
    pdf:
        fig-width: 6
        fig-height: 6
---


# Abundance Index Relationships  

This chapter will look at time series of abundance indices and the relationships between abundance index of different life stages.  

Relationships will only be examined in years where we have data from all life stages (using the variable `abundanceIndex_commonTimePeriod`).  

```{r}
library(tidyverse)
library(plotly)
library(corrplot)
library(GGally)
library(vegan)

source(here::here("functions.R"))

in_path <- here::here("data", "processed")
```

Read in data; trim to only common years; put life stages in proper order.  

```{r}
# brown
brn_all <- read.csv(here::here(in_path, "brn_abundAndEnv.csv"))
brn <- brn_all |> 
    filter(shrimp_year %in% common_shrimpYears) |> 
    mutate(life_stage = factor(life_stage, levels = ls_order_brn)) |> 
    rename(abundIndex = abundanceIndex_commonTimePeriod)
brn_wide <- brn |> 
    select(shrimp_year, life_stage, abundIndex) |> 
    pivot_wider(names_from = life_stage,
                values_from = abundIndex)

# white
wht_all <- read.csv(here::here(in_path, "wht_abundAndEnv.csv"))
wht <- wht_all |> 
    filter(shrimp_year %in% common_shrimpYears) |> 
    mutate(life_stage = factor(life_stage, levels = ls_order_wht)) |> 
    rename(abundIndex = abundanceIndex_commonTimePeriod)
wht_wide <- wht |> 
    select(shrimp_year, life_stage, abundIndex) |> 
    pivot_wider(names_from = life_stage,
                values_from = abundIndex)
```

Set up colors and shapes  

```{r}
# palcols <- RColorBrewer::brewer.pal(9, "PuBuGn")[2:9]
# palcols <- khroma::color("devon", reverse = TRUE)(12)[4:12]
palcols <- khroma::color("oslo", reverse = TRUE)(12)[3:11]
shps <- c(16, 15, 17, 9, 2, 1, 0, 5)
```


## Brown Shrimp  

### Time Series  

```{r}
#| label: fig-brnTimeSeries
#| fig-cap: "Time series of brown shrimp abundance index by life stage."
#| fig-width: 5
#| fig-height: 7

p <- ggplot(brn,
       aes(x = shrimp_year, y = abundIndex,
           col = life_stage,
           shape = life_stage)) +
    geom_hline(yintercept = 0,
               col = "gray40",
               linewidth = 0.5) +
    geom_point() +
    geom_line() +
    facet_wrap(~life_stage, scales = "free_y", ncol = 1) +
    theme_bw() +
    scale_color_manual(values = palcols) +
    scale_shape_manual(values = shps) +
    labs(title = "Brown Shrimp",
         x = "Shrimp Year",
         y = "Abundance Index",
         col = "Life Stage",
         shape = "Life Stage") +
    theme(legend.position = "none")

plotly_out(p)
```

```{r}
#| label: fig-brnBoxplotTimeSeries
#| fig-cap: "Time series of abundance index for brown shrimp, focused on the variability between life stages within each year."
#| fig-width: 7
#| fig-height: 5

p <- ggplot(brn,
            aes(x = shrimp_year, 
                y = abundIndex,
                group = shrimp_year
            )) +
    geom_hline(yintercept = 0,
               col = "gray40",
               linewidth = 0.5) +
    geom_boxplot(fill = "gray92",
                 col = "gray20",
                 alpha = 0.5,
                 outliers = FALSE) +
    geom_point(aes(col = life_stage,
                   shape = life_stage),
               size = 3) +
    theme_bw() +
    scale_color_manual(values = palcols) +
    scale_shape_manual(values = shps) +
    labs(title = "Brown shrimp abundance index by year",
         x = "Shrimp Year",
         y = "Abundance Index",
         col = "Life Stage",
         shape = "Life Stage") +
    theme(legend.position = "bottom")

plotly_out(p)
```


### PCA  

```{r}
brn_pca <- rda(brn_wide[2:ncol(brn_wide)])

# pull out info
eigs <- summary(eigenvals(brn_pca))
row.scores <- scores(brn_pca, choices = 1:3)$sites
col.scores <- data.frame(scores(brn_pca, choices = 1:3)$species) |> 
    rownames_to_column("column")
row.scores <- cbind(brn_wide, row.scores)
```


```{r}
#| label: tbl-brnPCAeigenvalues

knitr::kable(eigs,
             caption = "Variance explained by PC axes",
             digits = 4)

```

```{r}
#| label: fig-brnScreePlot
#| fig-cap: "Screeplot for brown shrimp PCA."

screeplot(brn_pca, type = "l")
```

```{r}
#| label: tbl-brnPCAloadings


knitr::kable(scores(brn_pca, 
                    choices = 1:3,
                    display = "species",
                    scaling = 0),
             caption = "unscaled loadings onto PC axes",
             digits = 4)
```

```{r}
#| label: fig-brnPCAbiplot
#| fig-cap: "Biplot from brown shrimp PCA. Each point represents a year. Columns were life stages, rows were years, and cell values were abundance index."

p <- ggplot(row.scores,
       aes(x = PC1,
           y = PC2)) +
    geom_segment(data = col.scores,
                 aes(x = 0, y = 0,
                     xend = PC1, yend = PC2),
                 col = "red",
                 arrow = arrow(length = unit(0.15, "inches")),
                 linewidth = 0.4,
                 alpha = 0.6) +
    geom_text(data = col.scores,
              aes(x = PC1*1.1, y = PC2*1.1,
                  label = column),
              col = "red",
              size = 3.5) +
    geom_text(aes(label = shrimp_year,
                  col = shrimp_year)) +
    khroma::scale_color_smoothrainbow() +
    theme_bw() +
    labs(title = "Brown Shrimp PCA biplot",
         x = paste0("PC1 (", round(eigs[2, 1]*100, 2), "%)"),
         y = paste0("PC2 (", round(eigs[2, 2]*100, 2), "%)"))

plotly_out(p)

```

```{r}
#| label: fig-brnPCA3d
#| fig-cap: "3-dimensional plot of brown shrimp PCA. Each point represents a year."
#| eval: false
#| eval-html: true

fig <- plot_ly() %>% 
    add_trace(data = row.scores,
              type = "scatter3d",
              mode = "markers",
              x = ~PC1,
              y = ~PC2,
              z = ~PC3,
              color = ~shrimp_year,
              marker = list(size = 6,
                            opacity = 0.9)
    ) 

# now add lines, text, and cross-hatches for each identified loading
for (i in 1:nrow(col.scores)) {
    fig <- fig %>% 
        add_paths(
            x = c(0, col.scores$PC1[i]),
            y = c(0, col.scores$PC2[i]),
            z = c(0, col.scores$PC3[i]),
            # type = "scatter3d",
            mode = "lines",
            line = list(color = "red", width = 4),
            showlegend = FALSE
        )  |>
        add_text(
            x = col.scores$PC1[i],
            y = col.scores$PC2[i],
            z = col.scores$PC3[i],
            text = col.scores$column[i],
            textfont = list(color = "red", size = 12),
            showlegend = FALSE
        )  %>% 
        add_trace(
            type = "scatter3d",
            mode = "markers",
            x = col.scores$PC1[i],
            y = col.scores$PC2[i],
            z = col.scores$PC3[i],
            marker = list(size = 6,
                          color = "red",
                          symbol = "cross"),
            showlegend = FALSE
        )
        
}

# print the plot
fig
```

### Correlations between life stages  

```{r}
#| label: fig-brnCorrMatrix
#| fig-cap: "Pairs plot of brown shrimp life stages, with Spearman correlation coefficients in the upper triangle."
brn_wide |> 
    select(-shrimp_year) |> 
    ggpairs(upper = list(continuous = wrap("cor", 
                                           method = "spearman")),
            lower = list(continuous = wrap(lowerFn)))
```

Juveniles positively correlated with subadults. Adult landings positively correlated with juveniles and subadults.  

### Cross-correlations (lags)  

The lag k value returned by ccf(x, y) estimates the correlation between x[t+k] and y[t].  

These are Pearson correlation coefficients, so lag-0 correlations will be a little different than the Spearman coefficients in the pairs plots.  

#### Juveniles vs. subadults

```{r}
#| label: fig-brnLagJvSubad
#| fig-cap: "Cross-correlation between juvenile and adult stages of brown shrimp"
print(ccf(brn_wide$juvenile, brn_wide$subadult))
```

The highest correlation is at lag 0: the time series are most correlated without lagging either of them. Lag -2 is almost as big, but doesn't seem to make much sense - juveniles 2 years ago correlated with this year's subadults? 

#### Postlarvae vs. adults  

Can we see if adults are correlated with the following year's postlarvae?  

```{r}
#| label: fig-brnLagPlvAd
#| fig-cap: "Cross-correlation between postlarval and adult stages of brown shrimp"

print(ccf(brn_wide$postlarval, brn_wide$adult))
```

Nothing shows up except something at 5 years, which could just be noise.  

## White Shrimp  

### Time Series  

```{r}
#| label: fig-whtTimeSeries
#| fig-cap: "Time series of white shrimp abundance index by life stage."
#| fig-width: 5
#| fig-height: 7

p <- ggplot(wht,
       aes(x = shrimp_year, y = abundIndex,
           col = life_stage,
           shape = life_stage)) +
    geom_hline(yintercept = 0,
               col = "gray40",
               linewidth = 0.5) +
    geom_point() +
    geom_line() +
    facet_wrap(~life_stage, scales = "free_y", ncol = 1) +
    theme_bw() +
    scale_color_manual(values = palcols) +
    scale_shape_manual(values = shps) +
    labs(title = "White Shrimp",
         x = "Shrimp Year",
         y = "Abundance Index",
         col = "Life Stage",
         shape = "Life Stage") +
    theme(legend.position = "none")

plotly_out(p)
```

```{r}
#| label: fig-whtBoxplotTimeSeries
#| fig-cap: "Time series of abundance index for white shrimp, focused on the variability between life stages within each year."
#| fig-width: 7
#| fig-height: 5

p <- ggplot(wht,
            aes(x = shrimp_year, 
                y = abundIndex,
                group = shrimp_year
            )) +
    geom_hline(yintercept = 0,
               col = "gray40",
               linewidth = 0.5) +
    geom_boxplot(fill = "gray92",
                 col = "gray20",
                 alpha = 0.5,
                 outliers = FALSE) +
    geom_point(aes(col = life_stage,
                   shape = life_stage),
               size = 3) +
    theme_bw() +
    scale_color_manual(values = palcols) +
    scale_shape_manual(values = shps) +
    labs(title = "White shrimp abundance index by year",
         x = "Shrimp Year",
         y = "Abundance Index",
         col = "Life Stage",
         shape = "Life Stage") +
    theme(legend.position = "bottom")

plotly_out(p)
```

### PCA  

```{r}
wht_pca <- rda(wht_wide[2:ncol(wht_wide)])

# pull out info
eigs <- summary(eigenvals(wht_pca))
row.scores <- scores(wht_pca, choices = 1:3)$sites
col.scores <- data.frame(scores(wht_pca, choices = 1:3)$species) |> 
    rownames_to_column("column")
row.scores <- cbind(wht_wide, row.scores)
```


```{r}
#| label: tbl-whtPCAeigenvalues

knitr::kable(eigs,
             caption = "Variance explained by PC axes",
             digits = 4)

```

```{r}
#| label: fig-whtScreePlot
#| fig-cap: "Screeplot for white shrimp PCA."

screeplot(wht_pca, type = "l")
```

```{r}
#| label: tbl-whtPCAloadings


knitr::kable(scores(wht_pca, 
                    choices = 1:3,
                    display = "species",
                    scaling = 0),
             caption = "unscaled loadings onto PC axes",
             digits = 4)
```

```{r}
#| label: fig-whtPCAbiplot
#| fig-cap: "Biplot from white shrimp PCA. Each point represents a year. Columns were life stages, rows were years, and cell values were abundance index."

p <- ggplot(row.scores,
       aes(x = PC1,
           y = PC2)) +
    geom_segment(data = col.scores,
                 aes(x = 0, y = 0,
                     xend = PC1, yend = PC2),
                 col = "red",
                 arrow = arrow(length = unit(0.15, "inches")),
                 linewidth = 0.4,
                 alpha = 0.6) +
    geom_text(data = col.scores,
              aes(x = PC1*1.1, y = PC2*1.1,
                  label = column),
              col = "red",
              size = 3.5) +
    geom_text(aes(label = shrimp_year,
                  col = shrimp_year)) +
    khroma::scale_color_smoothrainbow() +
    theme_bw() +
    labs(title = "White Shrimp PCA biplot",
         x = paste0("PC1 (", round(eigs[2, 1]*100, 2), "%)"),
         y = paste0("PC2 (", round(eigs[2, 2]*100, 2), "%)"))

plotly_out(p)

```

```{r}
#| label: fig-whtPCA3d
#| fig-cap: "3-dimensional plot of white shrimp PCA. Each point represents a year."
#| eval: false
#| eval-html: true

fig <- plot_ly() %>% 
    add_trace(data = row.scores,
              type = "scatter3d",
              mode = "markers",
              x = ~PC1,
              y = ~PC2,
              z = ~PC3,
              color = ~shrimp_year,
              marker = list(size = 6,
                            opacity = 0.9)
    ) 

# now add lines, text, and cross-hatches for each identified loading
for (i in 1:nrow(col.scores)) {
    fig <- fig %>% 
        add_paths(
            x = c(0, col.scores$PC1[i]),
            y = c(0, col.scores$PC2[i]),
            z = c(0, col.scores$PC3[i]),
            # type = "scatter3d",
            mode = "lines",
            line = list(color = "red", width = 4),
            showlegend = FALSE
        )  |>
        add_text(
            x = col.scores$PC1[i],
            y = col.scores$PC2[i],
            z = col.scores$PC3[i],
            text = col.scores$column[i],
            textfont = list(color = "red", size = 12),
            showlegend = FALSE
        )  %>% 
        add_trace(
            type = "scatter3d",
            mode = "markers",
            x = col.scores$PC1[i],
            y = col.scores$PC2[i],
            z = col.scores$PC3[i],
            marker = list(size = 6,
                          color = "red",
                          symbol = "cross"),
            showlegend = FALSE
        )
        
}

# print the plot
fig
```

### Correlations between life stages  

```{r}
#| label: fig-whtCorrMatrix
#| fig-cap: "Pairs plot of white shrimp life stages, with Spearman correlation coefficients in the upper triangle."
wht_wide |> 
    select(-shrimp_year) |> 
    ggpairs(upper = list(continuous = wrap("cor", 
                                           method = "spearman")),
            lower = list(continuous = wrap(lowerFn)))
```

Subadults positively correlated with landings (both spawning and non-spawning); interesting. Adult spawning from SEAMAP is positively correlated with adult spawning landings. Small negative correlation between non-spawning adults in SEAMAP vs. landings, though the graph mostly looks like scatter.  


### Cross-correlations (lags)  

The lag k value returned by ccf(x, y) estimates the correlation between x[t+k] and y[t].  

These are Pearson correlation coefficients, so lag-0 correlations will be a little different than the Spearman coefficients in the pairs plots.  

#### Juveniles vs. subadults

```{r}
#| label: fig-whtLagJvSubad
#| fig-cap: "Cross-correlation between juvenile and adult stages of white shrimp"

print(ccf(wht_wide$juvenile, wht_wide$subadult))
```

Nope, nothing to see.


#### Postlarvae vs. spawning adults  

Can we see if adults are correlated with the following year's postlarvae?  

```{r}
#| label: fig-whtLagPlvAdSp
#| fig-cap: "Cross-correlation between postlarval and adult (spawning) stages of white shrimp"

print(ccf(wht_wide$postlarval, wht_wide$adult.spawning))
```

Again, not much to see.   


## About

**Input `.qmd` file** for this section was: **`r xfun::with_ext(knitr::current_input(), "qmd")`** and can be found in the main directory.  

**Input data file(s)** - in `data/processed`:  

-   `brn_abundAndEnv.csv`
-   `wht_abundAndEnv.csv`
