---
format:
    html:
        fig-width: 5
        fig-height: 7
    pdf:
        fig-width: 6
        fig-height: 6
---


# Temperature thresholds  

This chapter will look at temperature thresholds and relationships to abundance indices.  

Relationships will only be examined in years where we have data from all life stages (using the variable `abundanceIndex_commonTimePeriod`).  

```{r}
library(tidyverse)
library(DescTools)   # for SpearmanRho
library(corrplot)    # for corrplot
library(plotly)

source(here::here("functions.R"))

in_path <- here::here("data", "processed")

```

Read in data; trim to only common years; put life stages in proper order.  

```{r}
# brown
brn_all <- read.csv(here::here(in_path, "brn_abundAndEnv.csv"))
brn <- brn_all |> 
    filter(shrimp_year %in% common_shrimpYears) |> 
    mutate(life_stage = factor(life_stage, levels = ls_order_brn)) |> 
    rename(abundIndex = abundanceIndex_commonTimePeriod)
brn_wide <- brn |> 
    select(shrimp_year, life_stage, abundIndex) |> 
    pivot_wider(names_from = life_stage,
                values_from = abundIndex)

# white
wht_all <- read.csv(here::here(in_path, "wht_abundAndEnv.csv"))
wht <- wht_all |> 
    filter(shrimp_year %in% common_shrimpYears) |> 
    mutate(life_stage = factor(life_stage, levels = ls_order_wht)) |> 
    rename(abundIndex = abundanceIndex_commonTimePeriod)
wht_wide <- wht |> 
    select(shrimp_year, life_stage, abundIndex) |> 
    pivot_wider(names_from = life_stage,
                values_from = abundIndex)
```

## Brown Shrimp  


```{r}
totalDays <- names(brn)[str_starts(names(brn), "totalDays")]
comp_grid <- expand.grid(life_stage = levels(brn$life_stage),
                         totalDays = totalDays)
comp_grid <- comp_grid |> 
    mutate(longestSpell = str_replace(totalDays, "totalDays", "longestSpell"))

comp_out <- list()

for(i in 1:nrow(comp_grid)){
    comp.stage <- comp_grid[i, "life_stage"]
    comp.totalDays <- comp_grid[i, "totalDays"]
    comp.longestSpell <- comp_grid[i, "longestSpell"]
    
    # subset the data frame
    comp.rows <- which(brn$life_stage == comp.stage)
    comp.cols <- which(names(brn) %in% c("shrimp_year", "abundIndex", as.character(comp.totalDays), as.character(comp.longestSpell)))
    
    
    comb <- brn[comp.rows, comp.cols]
    names(comb) <- c("shrimp_year", "abund_index", "totalDays", "longestSpell")
    
    # correlation and CI for total # days
    corr_totalDays <- SpearmanRho(comb$totalDays,
                                  comb$abund_index,
                                  use = "pairwise.complete.obs",
                                  conf.level = 0.95)
    corr_totalDays <- data.frame("totalDays_estimate" = corr_totalDays["rho"],
                                 "totalDays_ciLow" = corr_totalDays["lwr.ci"],
                                 "totalDays_ciHigh" = corr_totalDays["upr.ci"],
                                 row.names = NULL)
    
    # correlation and CI for longest # days
    corr_mostDays <- SpearmanRho(comb$longestSpell,
                                 comb$abund_index,
                                 use = "pairwise.complete.obs",
                                 conf.level = 0.95)
    corr_mostDays <- data.frame("consecDays_estimate" = corr_mostDays["rho"],
                                "consecDays_ciLow" = corr_mostDays["lwr.ci"],
                                "consecDays_ciHigh" = corr_mostDays["upr.ci"],
                                row.names = NULL)
    
    corr_out <- cbind(comp_grid[i, ],
                      corr_totalDays,
                      corr_mostDays)
    
    comp_out[[i]] <- corr_out
}

corrs_all <- bind_rows(comp_out) |> 
    mutate(totalDays_significant = case_when(totalDays_ciLow * totalDays_ciHigh > 0 ~ "yes",
                                            .default = "no"),
           consecDays_significant = case_when(consecDays_ciLow * consecDays_ciHigh > 0 ~ "yes",
                                            .default = "no"),
           thresholdC = as.numeric(str_remove(totalDays, "totalDays_lessThan")))

```




```{r}
#| label: fig-brnCorrTotalColdDays
#| fig-cap: "Estimated Spearman correlations between each life stage and the total number of winter days below a selected temperature threshold for brown shrimp. Points represent the estimated Spearman correlation coefficient, and vertical gray lines represnt the 95% confidence interval for the coefficient. Points are colored brown if the correlation is not significantly different from 0 (p > 0.05) and teal if the correlation is significant (p < 0.05). Significant points are also larger. These p-values are unadjusted, so expect around 1 'false' significant result for every 20 tests at this level."

p <- ggplot(corrs_all,
       aes(x = thresholdC,
           y = totalDays_estimate,
           color = totalDays_significant)) +
    geom_linerange(aes(ymin = totalDays_ciLow,
                       ymax = totalDays_ciHigh),
                   col = "gray60") +
    geom_point(aes(size = totalDays_significant),
               alpha = 0.9) +
    scale_size_manual(values = c(2, 3),
                      guide = "none") +
    scale_color_manual(values = c("#a6611a", "#018571")) +   # r color brewer BrBG, colorblind-safe, endpoints of 5-class palette
    geom_hline(yintercept = 0) +
    facet_wrap(~ life_stage, ncol = 1) +
    theme_bw() +
    labs(title = "Brown Shrimp vs. Total Cold Days",
         x = "Temperature threshold (C)",
         y = "Correlation estimate and 95% CI (Spearman's rho)",
         col = "Significant?")

plotly_out(p, legend_pos = "right")
```


```{r}
#| label: fig-brnCorrLongestColdSpell
#| fig-cap: "Estimated Spearman correlations between each life stage and the largest number of consecutive winter days below a selected temperature threshold for brown shrimp. Points represent the estimated Spearman correlation coefficient, and vertical gray lines represnt the 95% confidence interval for the coefficient. Points are colored brown if the correlation is not significantly different from 0 (p > 0.05) and teal if the correlation is significant (p < 0.05). These p-values are unadjusted, so expect around 1 'false' significant result for every 20 tests at this level."

p <- ggplot(corrs_all,
       aes(x = thresholdC,
           y = consecDays_estimate,
           color = consecDays_significant)) +
    geom_linerange(aes(ymin = consecDays_ciLow,
                       ymax = consecDays_ciHigh),
                   col = "gray60") +
    geom_point(aes(size = consecDays_significant),
               alpha = 0.9) +
    scale_size_manual(values = c(2, 3),
                      guide = "none") +
    scale_color_manual(values = c("#a6611a", "#018571")) +   # r color brewer BrBG, colorblind-safe, endpoints of 5-class palette
    geom_hline(yintercept = 0) +
    facet_wrap(~ life_stage, ncol = 1) +
    theme_bw() +
    labs(title = "Brown Shrimp vs. Longest Cold Spell",
         x = "Temperature threshold (C)",
         y = "Correlation estimate and 95% CI (Spearman's rho)",
         col = "Significant?")

plotly_out(p, legend_pos = "right")
```



## White Shrimp  


For white shrimp, not all life stages can be compared to the same winter, so they are broken up appropriately here.  

### Young (postlarvae + juvenile)  

```{r}
totalDays <- names(wht)[str_starts(names(wht), "postlarvWinter.totalDays")]

comp_grid <- expand.grid(life_stage = levels(wht$life_stage),
                         totalDays = totalDays)

comp_grid <- comp_grid |> 
    mutate(longestSpell = str_replace(totalDays, "totalDays", "longestSpell")) |> 
    filter(life_stage %in% c("postlarval", "juvenile"))

comp_out <- list()

for(i in 1:nrow(comp_grid)){
    comp.stage <- comp_grid[i, "life_stage"]
    comp.totalDays <- comp_grid[i, "totalDays"]
    comp.longestSpell <- comp_grid[i, "longestSpell"]
    
    # subset the data frame
    comp.rows <- which(wht$life_stage == comp.stage)
    comp.cols <- which(names(wht) %in% c("shrimp_year", "abundIndex", as.character(comp.totalDays), as.character(comp.longestSpell)))
    
    
    comb <- wht[comp.rows, comp.cols]
    names(comb) <- c("shrimp_year", "abund_index", "totalDays", "longestSpell")
    
    # correlation and CI for total # days
    corr_totalDays <- SpearmanRho(comb$totalDays,
                                  comb$abund_index,
                                  use = "pairwise.complete.obs",
                                  conf.level = 0.95)
    corr_totalDays <- data.frame("totalDays_estimate" = corr_totalDays["rho"],
                                 "totalDays_ciLow" = corr_totalDays["lwr.ci"],
                                 "totalDays_ciHigh" = corr_totalDays["upr.ci"],
                                 row.names = NULL)
    
    # correlation and CI for longest # days
    corr_mostDays <- SpearmanRho(comb$longestSpell,
                                 comb$abund_index,
                                 use = "pairwise.complete.obs",
                                 conf.level = 0.95)
    corr_mostDays <- data.frame("consecDays_estimate" = corr_mostDays["rho"],
                                "consecDays_ciLow" = corr_mostDays["lwr.ci"],
                                "consecDays_ciHigh" = corr_mostDays["upr.ci"],
                                row.names = NULL)
    
    corr_out <- cbind(comp_grid[i, ],
                      corr_totalDays,
                      corr_mostDays)
    
    comp_out[[i]] <- corr_out
}

corrs_all <- bind_rows(comp_out) |> 
    mutate(totalDays_significant = case_when(totalDays_ciLow * totalDays_ciHigh > 0 ~ "yes",
                                            .default = "no"),
           consecDays_significant = case_when(consecDays_ciLow * consecDays_ciHigh > 0 ~ "yes",
                                            .default = "no"),
           thresholdC = as.numeric(str_remove(totalDays, "postlarvWinter.totalDays_lessThan")))

```


```{r}
#| label: fig-whtYoungCorrTotalColdDays
#| fig-cap: "Estimated Spearman correlations between life stage and the total number of winter days below a selected temperature threshold for white shrimp in their postlarval/juvenile winters. Points represent the estimated Spearman correlation coefficient, and vertical gray lines represnt the 95% confidence interval for the coefficient. Points are colored brown if the correlation is not significantly different from 0 (p > 0.05) and teal if the correlation is significant (p < 0.05). Significant points are also larger. These p-values are unadjusted, so expect around 1 'false' significant result for every 20 tests at this level."
#| fig-height: 5

p <- ggplot(corrs_all,
       aes(x = thresholdC,
           y = totalDays_estimate,
           color = totalDays_significant)) +
    geom_linerange(aes(ymin = totalDays_ciLow,
                       ymax = totalDays_ciHigh),
                   col = "gray60") +
    geom_point(aes(size = totalDays_significant),
               alpha = 0.9) +
    scale_size_manual(values = c(2, 3),
                      guide = "none") +
    scale_color_manual(values = c("#a6611a", "#018571")) +   # r color brewer BrBG, colorblind-safe, endpoints of 5-class palette
    geom_hline(yintercept = 0) +
    facet_wrap(~ life_stage, ncol = 1) +
    theme_bw() +
    labs(title = "White Shrimp vs. Total Cold Days",
         x = "Temperature threshold (C)",
         y = "Correlation estimate and 95% CI (Spearman's rho)",
         col = "Significant?")

plotly_out(p, legend_pos = "right")
```

```{r}
#| label: fig-whtYoungCorrLongestColdSpell
#| fig-cap: "Estimated Spearman correlations between life stage and the largest number of consecutive winter days below a selected temperature threshold for white shrimp in their postlarval/juvenile winters. Points represent the estimated Spearman correlation coefficient, and vertical gray lines represnt the 95% confidence interval for the coefficient. Points are colored brown if the correlation is not significantly different from 0 (p > 0.05) and teal if the correlation is significant (p < 0.05). These p-values are unadjusted, so expect around 1 'false' significant result for every 20 tests at this level."
#| fig-height: 5

p <- ggplot(corrs_all,
       aes(x = thresholdC,
           y = consecDays_estimate,
           color = consecDays_significant)) +
    geom_linerange(aes(ymin = consecDays_ciLow,
                       ymax = consecDays_ciHigh),
                   col = "gray60") +
    geom_point(aes(size = consecDays_significant),
               alpha = 0.9) +
    scale_size_manual(values = c(2, 3),
                      guide = "none") +
    scale_color_manual(values = c("#a6611a", "#018571")) +   # r color brewer BrBG, colorblind-safe, endpoints of 5-class palette
    geom_hline(yintercept = 0) +
    facet_wrap(~ life_stage, ncol = 1) +
    theme_bw() +
    labs(title = "White Shrimp vs. Longest Cold Spell",
         x = "Temperature threshold (C)",
         y = "Correlation estimate and 95% CI (Spearman's rho)",
         col = "Significant?")

plotly_out(p, legend_pos = "right")
```


### Subadults and Adults  

```{r}
totalDays <- names(wht)[str_starts(names(wht), "subadultWinter.totalDays")]

comp_grid <- expand.grid(life_stage = levels(wht$life_stage),
                         totalDays = totalDays)

comp_grid <- comp_grid |> 
    mutate(longestSpell = str_replace(totalDays, "totalDays", "longestSpell")) |> 
    filter(!(life_stage %in% c("postlarval", "juvenile")))

comp_out <- list()

for(i in 1:nrow(comp_grid)){
    comp.stage <- comp_grid[i, "life_stage"]
    comp.totalDays <- comp_grid[i, "totalDays"]
    comp.longestSpell <- comp_grid[i, "longestSpell"]
    
    # subset the data frame
    comp.rows <- which(wht$life_stage == comp.stage)
    comp.cols <- which(names(wht) %in% c("shrimp_year", "abundIndex", as.character(comp.totalDays), as.character(comp.longestSpell)))
    
    
    comb <- wht[comp.rows, comp.cols]
    names(comb) <- c("shrimp_year", "abund_index", "totalDays", "longestSpell")
    
    # correlation and CI for total # days
    corr_totalDays <- SpearmanRho(comb$totalDays,
                                  comb$abund_index,
                                  use = "pairwise.complete.obs",
                                  conf.level = 0.95)
    corr_totalDays <- data.frame("totalDays_estimate" = corr_totalDays["rho"],
                                 "totalDays_ciLow" = corr_totalDays["lwr.ci"],
                                 "totalDays_ciHigh" = corr_totalDays["upr.ci"],
                                 row.names = NULL)
    
    # correlation and CI for longest # days
    corr_mostDays <- SpearmanRho(comb$longestSpell,
                                 comb$abund_index,
                                 use = "pairwise.complete.obs",
                                 conf.level = 0.95)
    corr_mostDays <- data.frame("consecDays_estimate" = corr_mostDays["rho"],
                                "consecDays_ciLow" = corr_mostDays["lwr.ci"],
                                "consecDays_ciHigh" = corr_mostDays["upr.ci"],
                                row.names = NULL)
    
    corr_out <- cbind(comp_grid[i, ],
                      corr_totalDays,
                      corr_mostDays)
    
    comp_out[[i]] <- corr_out
}

corrs_all <- bind_rows(comp_out) |> 
    mutate(totalDays_significant = case_when(totalDays_ciLow * totalDays_ciHigh > 0 ~ "yes",
                                            .default = "no"),
           consecDays_significant = case_when(consecDays_ciLow * consecDays_ciHigh > 0 ~ "yes",
                                            .default = "no"),
           thresholdC = as.numeric(str_remove(totalDays, "subadultWinter.totalDays_lessThan")))

```


```{r}
#| label: fig-whtAdultsCorrTotalColdDays
#| fig-cap: "Estimated Spearman correlations between life stage and the total number of winter days below a selected temperature threshold for white shrimp in their subadult and adult winters. Points represent the estimated Spearman correlation coefficient, and vertical gray lines represnt the 95% confidence interval for the coefficient. Points are colored brown if the correlation is not significantly different from 0 (p > 0.05) and teal if the correlation is significant (p < 0.05). Significant points are also larger. These p-values are unadjusted, so expect around 1 'false' significant result for every 20 tests at this level."


p <- ggplot(corrs_all,
       aes(x = thresholdC,
           y = totalDays_estimate,
           color = totalDays_significant)) +
    geom_linerange(aes(ymin = totalDays_ciLow,
                       ymax = totalDays_ciHigh),
                   col = "gray60") +
    geom_point(aes(size = totalDays_significant),
               alpha = 0.9) +
    scale_size_manual(values = c(2, 3),
                      guide = "none") +
    scale_color_manual(values = c("#a6611a", "#018571")) +   # r color brewer BrBG, colorblind-safe, endpoints of 5-class palette
    geom_hline(yintercept = 0) +
    facet_wrap(~ life_stage, ncol = 1) +
    theme_bw() +
    labs(title = "White Shrimp vs. Total Cold Days",
         x = "Temperature threshold (C)",
         y = "Correlation estimate and 95% CI (Spearman's rho)",
         col = "Significant?")

plotly_out(p, legend_pos = "right")
```


```{r}
#| label: fig-whtAdultsCorrLongestColdSpell
#| fig-cap: "Estimated Spearman correlations between life stage and the largest number of consecutive winter days below a selected temperature threshold for white shrimp in their subadult and adult winters. Points represent the estimated Spearman correlation coefficient, and vertical gray lines represnt the 95% confidence interval for the coefficient. Points are colored brown if the correlation is not significantly different from 0 (p > 0.05) and teal if the correlation is significant (p < 0.05). These p-values are unadjusted, so expect around 1 'false' significant result for every 20 tests at this level."

p <- ggplot(corrs_all,
       aes(x = thresholdC,
           y = consecDays_estimate,
           color = consecDays_significant)) +
    geom_linerange(aes(ymin = consecDays_ciLow,
                       ymax = consecDays_ciHigh),
                   col = "gray60") +
    geom_point(aes(size = consecDays_significant),
               alpha = 0.9) +
    scale_size_manual(values = c(2, 3),
                      guide = "none") +
    scale_color_manual(values = c("#a6611a", "#018571")) +   # r color brewer BrBG, colorblind-safe, endpoints of 5-class palette
    geom_hline(yintercept = 0) +
    facet_wrap(~ life_stage, ncol = 1) +
    theme_bw() +
    labs(title = "White Shrimp vs. Longest Cold Spell",
         x = "Temperature threshold (C)",
         y = "Correlation estimate and 95% CI (Spearman's rho)",
         col = "Significant?")

plotly_out(p, legend_pos = "right")
```


## About

**Input `.qmd` file** for this section was: **`r xfun::with_ext(knitr::current_input(), "qmd")`** and can be found in the main directory.  

**Input data file(s)** - in `data/processed`:  

-   `brn_abundAndEnv.csv`
-   `wht_abundAndEnv.csv`  
